# Data analysis of the data regularite-mensuelle-ter

## Loading packages
```{r}
library(tidyverse)
library(MASS)
library(caret)
library(ggplot2)
library(RColorBrewer)
library(corrplot)
library(klaR)
library(psych)
library(devtools)
library(patchwork)
library(zoo)
```

## Import data and setting graphic settings

```{r}
url = "https://ressources.data.sncf.com/api/explore/v2.1/catalog/datasets/regularite-mensuelle-ter/exports/csv?lang=fr&timezone=Europe%2FBerlin&use_labels=true&delimiter=%3B"

data = read.csv(url, sep = ";", header = TRUE)

mycol = brewer.pal(5, "Set1")
```

## Data preparation
```{r}
#Removing missing values
data = na.omit(data)

#Renaming col
colnames(data) = c("date", "region", "nbr_train_prog", "nbr_train_circ", "nbr_train_ann", "nbr_train_ret", "tx_reg", "prop", "comm")

#Making the region as factors
data$region = as.factor(data$region)

#Making the date variable as date in R
data$date = as.yearmon(data[,1], format = "%Y-%m")
```

## Data exploration
```{r}
#Overview of the data
head(data)
```

```{r}
#How much data per region?
table(data$region)
```

```{r}
#The boxplot for discrete variables
discrete_var = colnames(data)[3:6]

#We put the data in a form that is convenient for the boxplots
data_train = stack(data[,3:4])
data_train$ind = factor(data_train$ind, labels = c("Programmed trains", "Trains that ran"))

data_late = stack(data[,5:6])
data_late$ind = factor(data_late$ind, labels = c("Canceled trains", "Late trains"))

#The boxplots
ggplot(data_train, aes(x = ind, y = values, fill = ind)) +
  geom_boxplot(alpha=0.5) +
  theme_minimal() +
  labs(title = "Boxplots for the discrete variables",
       y = "Number of trains",
       x = "",
       fill = "Legend") 

ggplot(data_late, aes(x = ind, y = values, fill = ind)) +
  geom_boxplot(alpha=0.5) +
  theme_minimal() +
  labs(title = "Boxplots for the discrete variables",
       y = "Number of trains",
       x = "",
       fill = "Train status") 
```

```{r}
#Boxplot for the rates
data_rate = data["tx_reg"]
data_prop = data["prop"]

g1 = ggplot(data_rate, aes(y = tx_reg)) +
  geom_boxplot(alpha = 0.5, show.legend = FALSE, fill = "red") +
  labs(title="Boxplot for the regularity rate per month",
       y = "Rate") 

g2 = ggplot(data_prop, aes(y = prop)) +
geom_boxplot(alpha = 0.5, show.legend = FALSE, fill = "blue") +
labs(title="Boxplot for the number of on time\ntrains per one late",
     y = "Number of trains") 

g1 + g2
```

```{r}
#Regrouping the data to have numbers across all region per month
train_prog = data %>%
  group_by(date) %>%
  summarize(train_prog = sum(nbr_train_prog))

train_circ = data %>%
  group_by(date) %>%
  summarize(train_circ = sum(nbr_train_circ))
```

The graph
```{r}
#Points graph to visualize the difference between the trains that were programmed and the ones that ran
ggplot(train_prog, aes(x=date, y=train_prog)) + 
  geom_point(col = "blue") +
  geom_point(data = train_circ, aes(x=date,y=train_circ), col = "red", pch=4, alpha=0.5) +
  labs(title = "Number of trains programmed per month\nvs the ones that actually ran",
       x = "Date",
       y = "Number of trains") 
```
### For the PACA region 

For the following, we will be focusing in the PACA region
```{r}
#Filter vector
data_paca = data[data$region == "Provence Alpes Côte d'Azur",]
data_paca = data_paca[,-2] #Removing the region column
data_paca
```

It could be interesting to investigate the mean number of late trains per year
```{r}
#Summing the data for each year of late trains for the paca region
data_paca_year = data_paca %>%
  group_by(year = lubridate::year(date)) %>%
  summarize(nbr_train_prog = mean(nbr_train_prog), nrb_train_circ = mean(nbr_train_circ), nbr_train_ann = mean(nbr_train_ann), nbr_train_ret = mean(nbr_train_ret))
  
head(data_paca_year)
```


## Data analysis

In this next section, I thought it could be interesting to perform an ANOVA to test whether the differences between the late trains of the years 2021 up to 2024 are significant

To this end, we have to explicit our variables
IV : independent variable => the year, qualitative ordinal variable (4 modalities : 2021, 2022, 2023, 2024)
DV : dependent variable => number of late trains, quantitative discrete variable

Null hypothesis H0 : there are no difference between the number of late trains of the years 2021 up to 2024
Alternative hypothesis H1 : there is a difference between the number of late trains of the years 2021 up to 2024

Before diving into the test, we have to curate our data
```{r}
#Keeping only the date in our data using year function of the lubricate package
data_paca_aov = data.frame(date = year(data_paca$date), nbr_train_ret = data_paca$nbr_train_ret)
data_paca_aov = data_paca_aov[order(data_paca_aov$date),]
data_paca_aov = data_paca_aov[data_paca_aov$date>=2021,]
data_paca_aov
```

```{r}
#All is done, we can do the anova
anova = aov(data_paca_aov$nbr_train_ret~data_paca_aov$date)

summary(anova) #according to this, the difference is very strongly significant
```
Residuals independence
```{r}
#visual checking of the absence of correlation
plot(anova, 1) 
```

Normal distribution of the residuals
```{r}
#The residuals do not seem to be normally distributed (not following the line). Because the p value of the shapiro test is only a little below 0.05, we question the validity of our test
plot(anova, 2) 
shapiro.test(anova$residuals) 
```
Linear regression
```{r}
#Getting the linear regression from python, we want to know if the coefficients are significant to assess whether the regression is relevant or not
#=>As we can see, they are strongly significant : this regression might be a path to explore
linear_reg_paca = lm(data_paca$nbr_train_prog~data_paca$nbr_train_ret)
summary(linear_reg_paca)
```













